---
title: "RTSS Project Report"
subtitle: "Presentation of vaccine and disease interactions"
author: "Carl Murie and Raphael Gottardo"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
autosize: false
output: 
  ioslides_presentation:
    incremental: no
    keep_md: yes
    smaller: no
    
---
 
```{r overall-knitr-options, cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
library(knitr)
opts_chunk$set(cache=FALSE, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, results='asis')

library(Biobase)
library(data.table)
library(ggplot2)
library(gridExtra)
library(limma)
library(edgeR)
library(xtable)
library(GSEABase)
library(DT)
library(plyr)     ## join
library(stringr)  ## str_trim
library(heatmap3)

DOFIT <- FALSE ## if TRUE run lmFit, otherwise read it from file.
DOGSEA <- TRUE ## if FALSE read results from file
QUIET <- TRUE
ROOT <- "/shared/silo_researcher/Gottardo_R/cmurie_working/RTSS/RTSS_Full/"


dat <- read.table( paste0(ROOT, "data/allData.txt"), sep="\t")
anno <- read.csv( paste0(ROOT, "data/RTSSFullPheno.csv"))
meta <- read.table(paste0(ROOT, "data/allMeta.txt"), sep="\t", header=TRUE, row.names=1)

numberOfSubs <- length(unique(anno$SID))

## get plate locations of controls
controls <- setdiff(colnames(dat), anno$colID)

## separate controls from biological samples
conInd <- match(controls, colnames(dat))
conDat <- dat[,conInd]
bDat <- dat[,-conInd]
meta <- meta[,-conInd]

## make sure the phenotype and data files are ordered the same
ordInd <- match( colnames(bDat), anno$colID)
bPheno <- anno[ordInd,]
if(sum(colnames(bDat) != bPheno$colID) != 0) {stop("ERROR: column names of data don't match annotation file")}
if(sum(colnames(meta) != bPheno$colID) != 0) {stop("ERROR: column names of data don't match annotation file")}

numSub <- ncol(bDat)

 geneCut <- 20
 sampCut <- 75000
 pCut <- 0.01
 FDRCut <- 0.2
 
 ### remove low read samples
 totReads <- apply(bDat, 2, sum)
 dInd <- totReads <= sampCut
 deletes <- sum(dInd)
 dRan <- range(totReads[dInd])
  
 ## get idea of low to mid sample reads
 mInd <- totReads > sampCut & totReads <= 300000
 mids <- sum(mInd)
 mRan <- range(totReads[mInd])
 
  sampInd <- totReads > sampCut
  totReadsF <- uRan <- range(totReads[sampInd])
  fDat <- bDat[,sampInd]
  uMeta <- meta[,sampInd]
  
 ## filter genes with low read counts. Gene needs at least 20 samples with reads >= 5
 summ <- apply(fDat, 1, function(x) sum(x >= 5))
 summInd <- summ > geneCut
 
 ## before summing across technical replicates
 numGenes <- apply(bDat, 2, function(x) sum(x!=0))
 noZeros <- apply(bDat, 1, function(x) sum(x!=0))
 
 uDat <- fDat[summInd,]
 uPheno <- bPheno[sampInd,]
 if(sum(colnames(uDat) != uPheno$colID) != 0) {stop("ERROR: column names of data don't match annotation file")}
 
  ## get sample sizes by plate
  dataP=bPheno[bPheno$site=="BAGAMOYO",c(2,4,5,7,8,9)]
  total <- table(dataP$plate)
  bagSampBeforeP<- cbind(total, table(dataP[,c(1,4)]), table(dataP[,c(1,3)]),
                         table(dataP[,c(1,5)]),
                         table(dataP[,c(1,6)])[,-1])
  
  dataP=uPheno[uPheno$site=="BAGAMOYO",c(2,4,5,7,8,9)]
  total <- table(dataP$plate)
  bagSampAfterP<- cbind(total, table(dataP[,c(1,4)]), table(dataP[,c(1,3)]),
                         table(dataP[,c(1,5)]),
                         table(dataP[,c(1,6)])[,-1])

  dataP=bPheno[bPheno$site=="MANHICA",c(2,4,5,7,8,9)]
  total <- table(dataP$plate)
  manSampBeforeP<- cbind(total, table(dataP[,c(1,4)]), table(dataP[,c(1,3)]),
                         table(dataP[,c(1,5)]),
                         table(dataP[,c(1,6)])[,-1])
 
   dataP=uPheno[uPheno$site=="MANHICA",c(2,4,5,7,8,9)]
  total <- table(dataP$plate)
  manSampAfterP<- cbind(total, table(dataP[,c(1,4)]), table(dataP[,c(1,3)]),
                         table(dataP[,c(1,5)]),
                         table(dataP[,c(1,6)])[,-1])
  
  bagSampRemoveP <- bagSampBeforeP - bagSampAfterP
  manSampRemoveP <- manSampBeforeP - manSampAfterP
 
   ## Before Filtering: get sample sizes of design by location and stimulation
   data=bPheno[bPheno$site=="BAGAMOYO",c(4,5,7,8,9)]
   total <- table(data$stim)
   bagSampBefore <- cbind(total, table(data[,c(1,3)]), table(data[,c(1,2)]), table(data[,c(1,4)]),
              table(data[,c(1,5)])[,-1])[c(3,2,4,1),]
   data=bPheno[bPheno$site=="MANHICA",c(4,5,7,8,9)]
   total <- table(data$stim)
   manSampBefore <- cbind(total, table(data[,c(1,3)]), table(data[,c(1,2)]), table(data[,c(1,4)]),
              table(data[,c(1,5)])[,-1])[c(3,2,4,1),]
   
   ## After Filtering: get sample sizes of design by location and stimulation
   data=uPheno[uPheno$site=="BAGAMOYO",c(4,5,7,8,9)]
   total <- table(data$stim)
   bagSampAfter <- cbind(total, table(data[,c(1,3)]), table(data[,c(1,2)]), table(data[,c(1,4)]),
              table(data[,c(1,5)])[,-1])[c(3,2,4,1),]
   data=uPheno[uPheno$site=="MANHICA",c(4,5,7,8,9)]
   total <- table(data$stim)
   manSampAfter <- cbind(total, table(data[,c(1,3)]), table(data[,c(1,2)]), table(data[,c(1,4)]),
              table(data[,c(1,5)])[,-1])[c(3,2,4,1),]
   
   ## Get table showing which samples were removed during filtering
   bagSampRemove <- bagSampBefore - bagSampAfter
   manSampRemove <- manSampBefore - manSampAfter
   
  ## set up factors, relevel, rename
 uPheno$age <- factor(uPheno$agec, levels=unique(uPheno$agec), labels=c("old", "young"))
 uPheno$age  <- relevel(uPheno$age, "young")
 uPheno$agec <- NULL
 uPheno$stimulation <- relevel(uPheno$stimulation, "dmso")
 uPheno$vaccine <- relevel(uPheno$vaccine, "comparator")
 uPheno$case <- uPheno$rna_seq_casecon.m12
 uPheno$rna_seq_casecon.m12 <- NULL
 uPheno$case <- relevel(uPheno$case, "control")
 uPheno$match <- uPheno$rna_seq_match_id_m12
 uPheno$rna_seq_match_id_m12 <- NULL
 uPheno$TotalReads <- apply(uDat, 2, sum)
 
  ## fit basic analysis for QA
  myDes <- model.matrix(~stimulation*vaccine*age, data=uPheno)
  eDat <- ExpressionSet(assayData=as.matrix(uDat))
  normy <- calcNormFactors(eDat)
  libNorm <- colSums(exprs(eDat))*normy
   
```


## Experimental Design

Bulk RNASeq experiment assessing the effect of different stimulation on subjects treated with the RTSS malaria vaccine or comparator. There were a total of `r ncol(bDat)` samples.

**Subjects**: `r numberOfSubs` subjects.

**Vaccine**: rtss and comparator

**Stimulation**: dmso, csp, hbs

**Location**: bagamoyo and manhica

**Age**: 6-12 weeks (young) and 5-17 months (old).

**Outcome**: control (no malaria) and case (malaria)

**Visit**: M0 and M3 (initial visit and 3 months later)


##Analysis

Reads were aligned using BWA Aln version 0.7.10. The transcriptome was downloaded from
UCSC RefSeq (Human 19 or Mouse 10). Mitochondrial genes and ERCC reference sequences were
added. Sequence alignment and quantification were run at Broad Technology Labs. The data was voom transformed and a linear model was applied for statistical testing.

Samples with total read counts less than `r as.integer(sampCut)` were removed. Genes that did not have at least `r geneCut` samples (approximately 10% of the sample size) with read counts greater than 5 were removed. 

There were `r ncol(uDat)` samples and `r nrow(uDat)` genes remaining after filtering. 


##Sample Sizes: before sample filtering {.smaller}

There is a significant confound between age and location. Bagamoyo had samples only from the old age group and Manhica had mostly samples from the young age group.

<table border = 1>
<tr>
<td>
```{r results='asis', echo=FALSE}
   print(xtable(bagSampBefore, caption="Bagamoyo", align=rep("p", 10), auto=TRUE), size="small", type = "html")
```
</td>
</tr>
<tr>
<td>
```{r results='asis', echo=FALSE}
   print(xtable(manSampBefore, caption="Manhica", align=rep("p", 10), auto=TRUE), h.line=c(1:8), type = "html")
```
</td>
</tr>
</table>
        

##Sample Sizes: after sample filtering {.smaller}

<table border = 1>
<tr>
<td>
```{r results='asis', echo=FALSE}
   print(xtable(bagSampAfter, caption="Bagamoyo", align=rep("p", 10)), type = "html")
```
</td>
</tr>
<tr>
<td>
```{r results='asis', echo=FALSE}
   print(xtable(manSampAfter, caption="Manhica", align=rep("p", 10)), type = "html")
```
</td>
</tr>
</table>
           
       
##Samples lost due to filtering {.smaller}

There were a total of `r ncol(bDat) - ncol(uDat)` samples removed due to low read counts.

The Manhica location had the majority of discarded samples, particularly for the 5-17m group, even though it had fewer samples than Bagamoyo and very few 5-17m samples.

<table border = 1>
<tr>
<td>
```{r results='asis', echo=FALSE}
   print(xtable(bagSampRemove, caption="Bagamoyo", align=rep("p", 10)), type = "html")
```
</td>
</tr>
<tr>
<td>
```{r results='asis', echo=FALSE}
   print(xtable(manSampRemove, caption="Manhica", align=rep("p", 10)), type = "html")
```
</td>
</tr>
</table>

##Sample sizes by plate {.smaller}

There are no significant confounds by plate. Factors of biological interest are relatively balanced across plates.

```{r results='asis', echo=FALSE}
   print(xtable(bagSampBeforeP, caption="Bagamoyo", align=rep("p", 10)), type = "html")
```       

```{r results='asis', echo=FALSE}
   print(xtable(manSampBeforeP, caption="Manhica", align=rep("p", 10)), type = "html")
```   
   

##Samples lost by filtering by plate {.smaller}

The samples lost were mostly from Manhica and plate 3 in particular. 

```{r results='asis', echo=FALSE}
   print(xtable(bagSampRemoveP, caption="Bagamoyo", align=rep("p", 10)), type = "html")
```       

```{r results='asis', echo=FALSE}
   print(xtable(manSampRemoveP, caption="Manhica", align=rep("p", 10)), type = "html")
```

   

##Statistical Analysis {.smaller}

- Compare interaction effects between stimulants vs controls and vaccine (rtss vs comparator)
- Compare interaction effects between stimulants vs controls and disease (rtss vs comparator) separately for age groups (5-17m and 6-12w)
- Compare interaction effects of stimulants vs dmso and case vs control for subjects treated with rtss,s (comparator samples left out)
- Apply GSEA analysis (camera) to all comparisons using the Blood Transcriptional Modules (BTM)
- Apply individual gene testing (DEGs) to all comparisons 
- A Bayesian variance shrinkage estimator and the Benjamini-Hochberg FDR were applied.
- An FDR cutoff of `r FDRCut` was used for significance.
- Fold changes are calculated as stimulant - DMSO, rtss - comparator.


```{r gsea, results="asis"}
 
  ############################## GSEA analysis #####################################


   btmFile <- "/shared/silo_researcher/Gottardo_R/10_ref_files/GMT/BTM_for_GSEA_20131008.gmt"
   minGeneSetCut <- 5
   
    ## remove ama1 from study
    amaInd <- uPheno$stimulation == "ama1"
    uDat <- uDat[,!amaInd]
    uPheno <- uPheno[!amaInd,]
    libNorm <- libNorm[!amaInd]
    uPheno$stimulation <- as.factor(as.character(uPheno$stimulation))
    uPheno$stimulation <- relevel(uPheno$stimulation, "dmso")
 
   
   ## select only samples for M3 visit
   mInd <- uPheno$visit == "M3"
   uDat <- uDat[, mInd]
   uPheno <- uPheno[mInd,]
   libNorm <- libNorm[mInd]

  ######################### run stim vaccine analysis #####################################
   
  ## both young and old groups
  eDat <- ExpressionSet(assayData=as.matrix(uDat))
  pData(eDat) <- uPheno
  
  ## only young groups
  yInd <- uPheno$age=="young"
  yPheno <- uPheno[yInd,]
  ylibNorm <- libNorm[yInd]
  yDat <- ExpressionSet(assayData=as.matrix(uDat[,yInd]))
  pData(yDat) <- yPheno
     
  ## only old groups
  oInd <- uPheno$age=="old"
  oPheno <- uPheno[oInd,]
  olibNorm <- libNorm[oInd]
  oDat <- ExpressionSet(assayData=as.matrix(uDat[,oInd]))
  pData(oDat) <- oPheno
 
  myDes1 <- model.matrix(~plate + TotalReads + age + vaccine*stimulation, eDat)
  myDesy <- model.matrix(~plate + TotalReads + vaccine*stimulation, yDat)
  myDeso <- model.matrix(~plate + TotalReads + vaccine*stimulation, oDat)
 
  v <- voom(exprs(eDat), design=myDes1, plot=FALSE, lib.size=libNorm)
  vy <- voom(yDat, design=myDesy, plot=FALSE, lib.size=ylibNorm)
  vo <- voom(oDat, design=myDeso, plot=FALSE, lib.size=olibNorm)

  geneSet <- getGmt(btmFile)
  geneIds <- geneIds(geneSet)
  setsIndices <- ids2indices(geneIds, rownames(v$E))
  setsIndices <- setsIndices[sapply(setsIndices, length) > minGeneSetCut]
   
  coefs <- colnames(myDes1)[9:12]
  labs <- c("DMSO vs csp", "DMSO vs hbs", "DMSO vs csp:vaccine", "DMSO vs hbs:vaccine")
  agey <- c("both", "young", "old")
  
  ## assemble various fits
  allVoom <- list(v, vy, vo)
  allModels <- list(myDes1, myDesy, myDeso)

  if(DOGSEA) {
      
      ## apply GSEA to all models
      allGDT <- list()
      gMat <- NULL
      for(i in 1:length(allVoom)) {
          
          ageLab <- agey[i]
          v1 <- allVoom[[i]]
          mod <- allModels[[i]]
          
          tmpMat <- NULL
          coefInd <- 0
          for(c in coefs) {
            
              coefInd <- coefInd +1
              labb <- paste(labs[coefInd], ageLab, sep=": ")
              res <- camera(v1, setsIndices, design=mod, contrast=c, sort=TRUE)
              indo <- res$FDR <= FDRCut
              tmpMat <- c(tmpMat, sum(indo))
              
              if(sum(indo) > 0) {
                  allGDT[[labb]] <- cbind(res[indo, c("NGenes", "Direction")], 
                                              signif(res[indo, c("PValue","FDR")], 3))
              }
              else {
                allGDT[[labb]] <- -1
              }
              
          } ## end for c
             
          ## collect summary statistics
          gMat <- rbind(gMat, tmpMat)
          
      } ## end for i
   
       save(allGDT,  gMat, file=paste0(ROOT, "data/GSEAresLinNoAma.Rda"))
    
  } else {
       load(file=paste0(ROOT, "data/GSEAresLinNoAma.Rda"))
  }
  
  seqq <- c(1,5,9)
  newOrder <- names(allGDT)[c(seqq, seqq+1, seqq+2, seqq+3)]
  allGDT <- allGDT[newOrder]

  ## only show summary of interactions
  printGDT <- allGDT[7:12]
  printTab <- gMat[,3:4]
  dimnames(printTab) <- list(agey, c("csp", "hbs"))
  
 
  
  ###################### run case/control analysis ####################################
    
   ## select only case or control for match
   cInd <- (uPheno$case == "case" | uPheno$case == "control") & uPheno$vaccine=="rtss"
   libNormc <- libNorm[cInd]
   dDat <- uDat[,cInd]
   dPheno <- uPheno[cInd,]
   dPheno$case <- factor(dPheno$case, levels=c("control", "case"))
      
  ## both young and old groups
  eDat <- ExpressionSet(assayData=as.matrix(dDat))
  pData(eDat) <- dPheno
  
  ## only young groups
  yInd <- dPheno$age=="young"
  yPheno <- dPheno[yInd,]
  ylibNorm <- libNormc[yInd]
  yDat <- ExpressionSet(assayData=as.matrix(dDat[,yInd]))
  pData(yDat) <- yPheno
     
  ## only old groups
  oInd <- dPheno$age=="old"
  oPheno <- dPheno[oInd,]
  olibNorm <- libNormc[oInd]
  oDat <- ExpressionSet(assayData=as.matrix(dDat[,oInd]))
  pData(oDat) <- oPheno
 
  myDes1c <- model.matrix(~plate + TotalReads + age + case*stimulation, eDat)
  myDesyc <- model.matrix(~plate + TotalReads + case*stimulation, yDat)
  myDesoc <- model.matrix(~plate + TotalReads + case*stimulation, oDat)
 
  v <- voom(eDat, design=myDes1c, plot=FALSE, lib.size=libNormc)
  vy <- voom(yDat, design=myDesyc, plot=FALSE, lib.size=ylibNorm)
  vo <- voom(oDat, design=myDesoc, plot=FALSE, lib.size=olibNorm)
 
  deltaV <- v
  
  coefsc <- colnames(myDes1c)[9:12]
  labsc <- c("DMSO vs csp", "DMSO vs hbs", "DMSO vs csp:disease", "DMSO vs hbs:disease")
  agey <- c("both", "young", "old")
  
  ## assemble various fits
  allVoomc <- list(v, vy, vo)
  allModelsc <- list(myDes1c, myDesyc, myDesoc)
  
  ## apply GSEA to all models
  allGDTc <- allSelect <- list()
  gMatc <- NULL
  for(i in 1:length(allVoomc)) {
      
      ageLab <- agey[i]
      vc <- allVoomc[[i]]
      mod <- allModelsc[[i]]
      
      tmpMat <- NULL
      for(k in 1:length(coefsc)) {
         
          ## run camera on this comparison and data set
          c <- coefsc[k]
          labb <- paste(labsc[k], ageLab, sep=": ")
          res <- camera(vc, setsIndices, design=mod, contrast=c, sort=TRUE)
          indo <- res$FDR <= FDRCut
          tmpMat <- c(tmpMat, sum(indo))
          
          if(sum(indo) > 0) {
              allGDTc[[labb]] <- cbind(res[indo, c("NGenes", "Direction")], 
                                          signif(res[indo, c("PValue","FDR")], 3))
          } else {
            allGDTc[[labb]] <- -1
          }
          
          ## get index label for vaccine model
          selectC <- coefs[k]
          selectLab <-  paste(labs[k], ageLab, sep=": ")
          selectDT <- allGDT[[selectLab]]
          
          ## get disease results for significant gene sets with vaccine model
          if(length(selectDT) == 1) {
             allSelect[[labb]] <- -1            
          } else {
             tmp2 <- cbind(res[rownames(selectDT), c("NGenes", "Direction")], 
                          signif(res[rownames(selectDT), c("PValue","FDR")], 3))
             
             ## recalculate FDRs on just the previously selected gene sets
             fdrs <- signif(p.adjust(tmp2$PValue, method="BH"), 3)
             tmp2$FDR <- fdrs
             tmp2 <- tmp2[order(tmp2$PValue),]
             allSelect[[labb]] <- tmp2
          }
          
      } ## end for k
         
      ## collect summary statistics
      gMatc <- rbind(gMatc, tmpMat)
      
  } ## end for v

  seqq <- c(1,5,9)
  newOrder <- names(allGDTc)[c(seqq, seqq+1, seqq+2, seqq+3)]
  allGDTc <- allGDTc[newOrder]
  allSelect <- allSelect[newOrder]

  ## only show summary of interactions
  printGDTc <- allGDTc[7:12]
  printSelect <- allSelect[7:12]
  printTabc <- gMatc[,3:4]
  dimnames(printTabc) <- list(agey, c("csp", "hbs"))
  

  ## generate output string for knit_child
  outcase = NULL
  for (i in 1:length(printGDT)) {
    
    ## vaccine model
    knit_expanded <- paste0("##`r names(printGDT)[",i, "]` {.smaller}\n")
    outcase = c(outcase, knit_expanded)
    
     if(is.null(dim(printGDT[[i]]))) {
           knit_expanded <- "No significant gene sets found\n" 
      }else {
        knit_expanded <-  paste0("`r datatable(printGDT[[", i, "]])`\n")
      }
    outcase = c(outcase, knit_expanded)
    
    ## disease model for significant vaccine effects
    knit_expanded <- paste0("##`r names(printSelect)[",i, "]`| disease model for significant vaccine effects {.smaller}")
    outcase = c(outcase, knit_expanded)
    
     if(is.null(dim(printSelect[[i]]))) {
           knit_expanded <- "No significant gene sets found\n" 
      }else {
        knit_expanded <-  paste0("`r datatable(printSelect[[", i, "]])`\n")
      }
    outcase = c(outcase, knit_expanded)
    
    ## disease model
    knit_expanded <- paste0("##`r names(printGDTc)[",i, "]` {.smaller}\n")
    outcase = c(outcase, knit_expanded)
    
     if(is.null(dim(printGDTc[[i]]))) {
           knit_expanded <- "No significant gene sets found\n" 
      }else {
        knit_expanded <-  paste0("`r datatable(printGDTc[[", i, "]])`\n") 
      }
    outcase = c(outcase, knit_expanded)
    
  } ## end for i
  
  
  
  ##################################################################################
  ## code to calculate deltas and provide some specific graphs. Not part of report.
  ###################################################################################
  if(FALSE) {
    
  ## crudely calculate stim - dmso for each subject TODO - figure out how to do this with datatable
  pheno <- deltaPheno## eDat generated in line 331-334 (only M3 and no ama1)
  ##dat <- exprs(eDat)
  dat <- deltaV$E
  
  
  
  delta <- stimFac <- caseFac <- ageFac <- vaccineFac <- NULL
  for(s in unique(pheno$SID)) {
     for(stim in c("csp", "hbs")) {
         indStim <- pheno$SID==s & pheno$stimulation==stim
         indDMSO <- pheno$SID==s & pheno$stimulation=="dmso"
         if(sum(indStim == 1) & sum(indDMSO == 1)) {
             if(pheno$case[indStim] != pheno$case[indDMSO]) {stop("ERROR case")}
             if(pheno$age[indStim] != pheno$age[indDMSO]) {stop("ERROR age")}
             delta <- cbind(delta, dat[,indStim] - dat[,indDMSO])
             stimFac <- c(stimFac, stim)
             caseFac <- c(caseFac, as.character(pheno$case[indStim]))
             ageFac <- c(ageFac, as.character(pheno$age[indStim]))
             vaccineFac <- c(vaccineFac, as.character(pheno$vaccine[indStim]))
         } ## end for sum
     } ## end for stim
  } ## end for s
  
  ### vaccine model
  
  vLabs <- c("DMSO vs csp:vaccine: both", "DMSO vs hbs:vaccine: both")
  vaccineFac[vaccineFac=="comparator"] <- "comp"
  vaccineFac <- factor(vaccineFac, levels=c("rtss", "comp"))
  ageFac <- factor(ageFac, levels=c("young", "old"))  
 
  ncolDat <- rbind(c(6,4), c(6,3), c(6,4), c(4,3), c(6,4), c(6,3), c(4,3))
 
  ggPlotListV <- list()
  for(set in rownames(allGDT[[vLabs[2]]])) {

     sigGenes <- geneIds[[set]]  ## get gene symbols for this set
     sigDat <- delta[intersect(sigGenes, rownames(delta)),]
   
     ## look at means of each level out of interest
     mRTSS <- apply(sigDat[,vaccineFac=="rtss"], 2, mean)
     mComp <- apply(sigDat[,vaccineFac=="comp"], 2, mean)
     print(allGDT[[vLabs[2]]][set,])
     print(mean(mRTSS) - mean(mComp))
  
     ## melt age factor  
     colnames(sigDat) <- ageFac
     dt1 <- melt(sigDat)
 
     ## melt case factor  ? is there a better way to melt the two factors ?
     colnames(sigDat) <- vaccineFac
     dt <- melt(sigDat)
     colnames(dt) <- c("gene", "vaccine", "expression")
     dt$age <- factor(dt1$Var2, levels=c("young", "old"), ordered=TRUE)
  
     ggPlotListV[[set]] <- ggplot(data=dt, aes(x=vaccine, y=expression)) + 
                                 geom_boxplot(outlier.size=-1) +
       geom_point(aes(color=age),position=position_jitterdodge(), alpha=0.2, size=0.3) +
        facet_wrap(~gene, scales= "free", ncol=6) + ggtitle(set) +theme(aspect.ratio=1)
    
  } ## end for set
  
  pdf(file=paste(ROOT, "plots/boxPlotsVaccineHBS.pdf", sep=""))
  print(ggPlotListV[[1]]) 
  print(ggPlotListV[[2]]) 
  print(ggPlotListV[[3]])
  print(ggPlotListV[[4]])
  print(ggPlotListV[[5]])
  print(ggPlotListV[[6]])
  print(ggPlotListV[[7]])
  dev.off() 
  
 

  #### disease model
  dInd <- (caseFac == "case" | caseFac == "control") & vaccineFac=="rtss" & stimFac=="hbs"
  deltaCase <- delta[,dInd]
  caseCase <- caseFac[dInd]
  ageCase <- ageFac[dInd]
  stimCase <- stimFac[dInd]

  ## shorten control level for graph
  caseCase[caseCase=="control"] <- "con"
  
  ncolDat <- rbind(c(6,4), c(3,3), c(6,4))
  
  ## comparisons of interest (allGDTc) only second comparison has significant results
  comps <- c("DMSO vs csp:disease: both", "DMSO vs hbs:disease: both")
  
  ggPlotList <- list()
  colInd <- 0
  for(set in rownames(allGDTc[[comps[2]]])) {

     colInd <- colInd + 1
     sigGenes <- geneIds[[set]]  ## get gene symbols for this set
     sigDat <- deltaCase[intersect(sigGenes, rownames(deltaCase)),]
   
     ## look at means of each level out of interest
     mCase <- apply(sigDat[,caseCase=="case"], 2, mean)
     mCon <- apply(sigDat[,caseCase=="control"], 2, mean)
     ##print(mean(mCase) - mean(mCon))
  
     ## melt age factor  
     colnames(sigDat) <- ageCase
     dt1 <- melt(sigDat)
 
     ## melt case factor  ? is there a better way to melt the two factors ?
     colnames(sigDat) <- caseCase
     dt <- melt(sigDat)
     colnames(dt) <- c("gene", "disease", "expression")
     dt$age <- factor(dt1$Var2, levels=c("young", "old"), ordered=TRUE)
     dt$disease <- factor(dt$disease, levels=c("case", "con"), ordered=TRUE)
 
     ggPlotList[[set]] <- ggplot(data=dt, aes(x=disease, y=expression)) + 
                                 geom_boxplot(outlier.size=-1) +
       geom_point(aes(color=age),position=position_jitterdodge(), alpha=0.4, size=0.65) +
        ##facet_wrap(~gene, scales= "free", ncol=ncolDat[colInd,1], nrow=ncolDat[colInd,2]) +
        facet_wrap(~gene, scales= "free", ncol=6, nrow=4) +
       ggtitle(set) + theme(aspect.ratio = 1)
    
  } ## end for set
  
 
  
  pdf(file=paste(ROOT, "plots/boxPlotsDiseaseHBS.pdf", sep=""))
  print(ggPlotList[[1]]) 
  print(ggPlotList[[2]]) 
  print(ggPlotList[[3]])
  dev.off() 
  
  
  #### disease model csp
  dIndh <- (caseFac == "case" | caseFac == "control") & vaccineFac=="rtss" & stimFac=="csp"
  deltaCaseh <- delta[,dIndh]
  caseCaseh <- caseFac[dIndh]
  ageCaseh <- ageFac[dIndh]
  stimCaseh <- stimFac[dIndh]

  
  ## put IL2 plots together
  
  ILvac <- delta["IL2",]
  ILVacdf <- data.frame(expression=ILvac, vaccine=vaccineFac, age=ageFac, stim=stimFac)
  ILVaccsp <- ILVacdf[stimFac=="csp",]
  ILVachbs <- ILVacdf[stimFac=="hbs",]
 
  
  ## hbs
  ILdis <- deltaCase["IL2",]
  ILDisdf <- data.frame(expression=ILdis, disease=caseCase, age=ageCase, stim=stimCase)
  ILDisdf$disease <- factor(ILDisdf$disease, levels=c("control", "case"))

  ## csp
  ILdish <- deltaCaseh["IL2",]
  ILDisdfh <- data.frame(expression=ILdish, disease=caseCaseh, age=ageCaseh, stim=stimCaseh)
  ILDisdfh$disease <- factor(ILDisdfh$disease, levels=c("control", "case"))

  
  ## csp plots
  ggplot(data=ILVaccsp, aes(x=vaccine, y=expression)) + 
                                 geom_boxplot(outlier.size=-1) +
       geom_point(aes(color=age),position=position_jitterdodge(), alpha=1, size=1) +
       ggtitle("IL2: vaccine model")
    
  ggplot(data=ILDisdfh, aes(x=disease, y=expression)) + 
                                 geom_boxplot(outlier.size=-1) +
       geom_point(aes(color=age),position=position_jitterdodge(), alpha=1, size=1) +
       ggtitle("IL2: disease model")

  
  
  ## hbs plots
  pdf(file=paste0(ROOT, "plots/boxPlotsIL2.pdf"))
  ggplot(data=ILVachbs, aes(x=vaccine, y=expression)) + 
                                 geom_boxplot(outlier.size=-1) +
       geom_point(aes(color=age),position=position_jitterdodge(), alpha=1, size=1) +
       ggtitle("IL2: vaccine model")
    
  ggplot(data=ILDisdf, aes(x=disease, y=expression)) + 
                                 geom_boxplot(outlier.size=-1) +
       geom_point(aes(color=age),position=position_jitterdodge(), alpha=1, size=1) +
       ggtitle("IL2: disease model")
  dev.off() 
  
  
   pdf(file=paste(ROOT, "plots/boxPlotsDiseaseSig.pdf", sep=""))

  ggplot(data=dtm81both, aes(x=disease, y=expression)) + geom_boxplot(outlier.size = -1) + 
    geom_jitter(alpha=0.2) + facet_wrap(~gene, scales= "free") +ggtitle(sigSetM81)
  ggplot(data=dtm81old, aes(x=disease, y=expression)) + geom_boxplot(outlier.size = -1) + 
    geom_jitter(alpha=0.2) + facet_wrap(~gene, scales= "free") +ggtitle(sigSetM81)
  ggplot(data=dtm32old, aes(x=disease, y=expression)) + geom_boxplot(outlier.size = -1) + 
    geom_jitter(alpha=0.2) + facet_wrap(~gene, scales= "free") +ggtitle(sigSetM32)
  
   ggplot(data=dtm81both, aes(x=disease, y=expression)) + geom_violin(outlier.size = -1) + 
    geom_jitter(alpha=0.2) + facet_wrap(~gene, scales= "free") +ggtitle(sigSetM81)
  ggplot(data=dtm81old, aes(x=disease, y=expression)) + geom_violin(outlier.size = -1) + 
    geom_jitter(alpha=0.2) + facet_wrap(~gene, scales= "free") +ggtitle(sigSetM81)
  ggplot(data=dtm32old, aes(x=disease, y=expression)) + geom_violin(outlier.size = -1) + 
    geom_jitter(alpha=0.2) + facet_wrap(~gene, scales= "free") +ggtitle(sigSetM32)
  
 dev.off()
  
 
 ## look at vaccine model
  sigSet <- sigSetM81 <- "enriched in myeloid cells and monocytes (M81)"
  sigGenes <- geneIds[[sigSet]]
  
  sigSet <- sigSetM32 <- "CORO1A-DEF6 network (I) (M32.2)"
  sigGenes <- geneIds[[sigSet]]
  
  
  ##stimInd <- stimFac=="hbs" ## M81 both
  stimInd <- stimFac=="hbs" & ageFac=="old"  ## M81 old 
  
  sigDat <- delta[intersect(sigGenes, rownames(delta)), stimInd]
  casey <- caseFac[stimInd]

  sigDatCase <- sigDat[,casey=="case"]
  sigDatCon <- sigDat[,casey=="control"]
  mCase <- apply(sigDatCase, 2, mean)
  mCon <- apply(sigDatCon, 2, mean)
  sigDatAll <- cbind(sigDatCase[,order(mCase, decreasing=TRUE)], 
                     sigDatCon[, order(mCon, decreasing=TRUE)])
  caseAll <- matrix(c(rep(1, ncol(sigDatCase)), rep(2, ncol(sigDatCon))), ncol=1)
  caseName <- matrix(c(rep("case", ncol(sigDatCase)), rep("control", ncol(sigDatCon))), ncol=1)
  colnames(caseAll) <- "case"
  
  colnames(sigDatAll) <- caseName
  dt <- melt(sigDatAll)
  colnames(dt) <- c("gene", "disease", "expression")
 
  dtm81both <- dt
  dtm81old <- dt
  dtm32old <- dt
 
 
 
 
 
 
  pdf(file=paste(ROOT, "plots/heatMapsM81hbsOld.pdf", sep=""))
  heatmap3(sigDatAll, Rowv=TRUE, Colv=NA, ColSideColors=caseAll, main=sigSet)
  dev.off()
  
  for(n in 1:length(allGDTc)) {
     namey <- names(allGDTc)[n]
     tmp <- allGDTc[[n]]
     if(length(tmp)==1) {
        next()
     }
     stimm <- stimInd[n]
     agey <- ageInd[n]
     
     for(i in rownames(tmp)) {
        sigGenes <- geneIds[[i]]
        if(n==1 | n==4 | n==7 | n==10) { ## use both age groups
           sigDat <- delta[intersect(sigGenes, rownames(delta)), stimFac==stimm]
           caseCol <-  matrix(as.numeric(as.factor(caseFac[stimFac==stimm])), ncol=1)
        } else {                         ## use either young or old age groups
           sigDat <- delta[intersect(sigGenes, rownames(delta)), stimFac==stimm & ageFac==agey]
           caseCol <-  matrix(as.numeric(as.factor(caseFac[stimFac==stimm & ageFac==agey])), ncol=1)
        }
        colnames(caseCol) <- "case"
        heatmap3(sigDat, Rowv=TRUE, Colv=TRUE, ColSideColors=caseCol, main=paste(namey, i, sep=": "))
     } ## end for i
     
    
  } ## end for n
  dev.off()
  } ## end if FALSE
  ###################################################################################
  ################# end of delta section ############################################
  ###################################################################################
  
```

<!-- knit GSEA results--> 
`r paste(knit_child(text = outcase), collapse = '\n')`



##GSEA summary stimulation:vaccine interaction| FDR cutoff: `r FDRCut`

Number of significant BTM gene sets for dmso versus stimulants and vaccine interaction.


```{r results='asis', echo=FALSE}
   print(xtable(printTab, align=rep("c", ncol(printTab)+1), auto=TRUE), type = "html")
```


##GSEA summary stimulation:disease interaction| FDR cutoff: `r FDRCut`

Number of significant BTM gene sets for dmso versus stimulants and disease nteraction.


```{r results='asis', echo=FALSE}
   print(xtable(printTabc, align=rep("c", ncol(printTabc)+1), auto=TRUE), type = "html")
```



```{r voom, cache=FALSE}
  
   ####################### stimulation vaccine effects ################################

  ## both young and old groups
  eDat <- ExpressionSet(assayData=as.matrix(uDat))
  pData(eDat) <- uPheno
  
  ## only young groups
  yInd <- uPheno$age=="young"
  yPheno <- uPheno[yInd,]
  ylibNorm <- libNorm[yInd]
  yDat <- ExpressionSet(assayData=as.matrix(uDat[,yInd]))
  pData(yDat) <- yPheno
     
  ## only old groups
  oInd <- uPheno$age=="old"
  oPheno <- uPheno[oInd,]
  olibNorm <- libNorm[oInd]
  oDat <- ExpressionSet(assayData=as.matrix(uDat[,oInd]))
  pData(oDat) <- oPheno
 
  
  myDes1 <- model.matrix(~plate + TotalReads + age + vaccine*stimulation, eDat)
  myDesy <- model.matrix(~plate + TotalReads + vaccine*stimulation, yDat)
  myDeso <- model.matrix(~plate + TotalReads + vaccine*stimulation, oDat)
   
  v <- voom(exprs(eDat), design=myDes1, plot=FALSE, lib.size=libNorm)
  vy <- voom(yDat, design=myDesy, plot=FALSE, lib.size=ylibNorm)
  vo <- voom(oDat, design=myDeso, plot=FALSE, lib.size=olibNorm)
    
  ## with the large matrix fitting the model takes 2-3 hrs. If it is already calculated
  ## just read in the fitted model.
  if(FALSE) {
    
      ## age == both model
      system.time(ranCor <- duplicateCorrelation(v, design=myDes1,
                                     block=uPheno$SID)$consensus.correlation)
     fit1 <- lmFit(v, myDes1, block=uPheno$SID, correlation=ranCor)
     save(fit1, ranCor, file=paste0(ROOT, "data/lmFitDataWithCoVarM3NoAMA.Rda"))
     
     ## run young age model
     ranCorYoung <- duplicateCorrelation(vy, design=myDesy,
                                     block=yPheno$SID)$consensus.correlation
     fitYoung <- lmFit(vy, myDesy, block=yPheno$SID, correlation=ranCorYoung)
     save(fitYoung, ranCorYoung, file=paste0(ROOT, "data/lmFitDataWithCoVarYoungNoAMA.Rda"))
   
     ## run old age model
     ranCorOld <- duplicateCorrelation(vo, design=myDeso,
                                     block=oPheno$SID)$consensus.correlation
     fitOld <- lmFit(vo, myDeso, block=oPheno$SID, correlation=ranCorOld)
     save(fitOld, ranCorOld, file=paste0(ROOT, "data/lmFitDataWithCoVarOldNoAMA.Rda"))
     
  } else {
     load(file=paste0(ROOT, "data/lmFitDataWithCoVarM3NoAMA.Rda"))    ## load fit1
     load(file=paste0(ROOT, "data/lmFitDataWithCoVarYoungNoAMA.Rda")) ## load fitYoung
     load(file=paste0(ROOT, "data/lmFitDataWithCoVarOldNoAMA.Rda"))   ## load fitOld
  }
  
  coefs <- colnames(myDes1)[9:12]
  labs <- c("DMSO vs csp", "DMSO vs hbs", "DMSO vs csp:vaccine","DMSO vs hbs:vaccine")
  agey <- c("both", "young", "old")
  
  
  ## assemble various fits
  allFit <- list(fit1, fitYoung, fitOld)
  
  ## extract data.tables and graphs from all fits
  allDT <- allDEG <- allGraphs <- list()
  ageInd <- 0
  for(fit in allFit) {
      
      ## set age for this fit
      ageInd <- ageInd + 1
      ageLab <- agey[ageInd]
      
      ## shrinkage
      fit2 <- eBayes(fit, trend=FALSE)
    
      ## cycle through coefficients
      colInd <- c(1,4,5)
      DEG <- NULL
      coefInd <- 0
      for(c in coefs) {
           
           coefInd <- coefInd+1
           labb <- paste(labs[coefInd], ageLab, sep=": ")
    
           ## generate datatables
           tmp <- topTable(fit2, number=Inf, coef=c, sort="P")
           ##write.csv(tmp, file=paste0(ROOT, "data/", labs[i], "Result.csv"), row.names=TRUE)
           indy <- tmp$P.Value <= pCut
           tmpy <-  signif(tmp[indy, colInd], 3)
           colnames(tmpy) <- c("log FC",  "P.value", "FDR")
           allDT[[labb]] <- tmpy
           DEG <- c(DEG, sum(tmpy$FDR <= FDRCut))
   
           ## generate graphs
          graphSet <- list()
          ress1 <- tmp
          ress1$FDR <- ifelse(ress1$adj.P.Val > FDRCut, "NonSig", "Sig")   ## label significant genes
          
          graphSet[[1]] <- ggplot(ress1, aes(x=P.Value)) + geom_histogram(binwidth=.05) +
            xlab("p-value") + ggtitle(label=labb) + theme_bw() + theme(aspect.ratio=1,
                                                                       axis.text.x=element_text(size=10),
                                                plot.title = element_text(size = 15)) + 
            scale_x_continuous(breaks=c(0, 0.5, 1))
          
          graphSet[[2]] <- ggplot(ress1, aes(x=logFC, y=-log10(P.Value), color=FDR)) +
            geom_point(shape=1) + ggtitle("logFC vs -log10 p-value") + theme_bw() + 
            theme(axis.title.y=element_text(size=10), axis.text.x=element_text(size=6), 
                  plot.title = element_text(size = 13)) + 
            geom_vline(xintercept=c(-1,1), linetype="dashed", color="blue") + 
          geom_hline(yintercept=-log10(c(0.01)), linetype="dashed", color="blue") +
            theme(legend.position="none") +  scale_colour_manual(values=c("black", "red")) +
            theme(aspect.ratio=1)
          
          graphSet[[3]] <- ggplot(ress1, aes(x=adj.P.Val)) + geom_histogram(binwidth=.05) +
            xlab("p-value") + ggtitle("FDR") + theme_bw() + 
            theme(aspect.ratio=1, plot.title = element_text(size = 15)) + 
            scale_x_continuous(breaks=c(0, 0.5, 1))
          
          allGraphs[[labb]] <- graphSet
      }
      
      ##allDT[[ageLab]] <- DT
      ##allGraph[[ageLab]] <- graphs
      allDEG <- rbind(allDEG, DEG)
   
  } ## end for fit
  
  ## reorder the lists for better presentation
  seqq <- c(1,5,9)
  newOrder <- names(allGraphs)[c(seqq, seqq+1, seqq+2, seqq+3)]
  allGraphs <- allGraphs[newOrder]
  allCompDT <- allDT
  allDT <- allDT[newOrder]
  
  ## show only results for interactions
  printGraphs <- allGraphs[7:12]
  printDT <- allDT[7:12]
  printDEGTab <- allDEG[,3:4]
  dimnames(printDEGTab) <- list(agey, c("csp", "hbs"))

  ##################### stimulation disease analysis ##################################
  
    ## both young and old groups
  eDat <- ExpressionSet(assayData=as.matrix(dDat))
  pData(eDat) <- dPheno
  
  ## only young groups
  yInd <- dPheno$age=="young"
  yPheno <- dPheno[yInd,]
  ylibNorm <- libNormc[yInd]
  yDat <- ExpressionSet(assayData=as.matrix(dDat[,yInd]))
  pData(yDat) <- yPheno
     
  ## only old groups
  oInd <- dPheno$age=="old"
  oPheno <- dPheno[oInd,]
  olibNorm <- libNormc[oInd]
  oDat <- ExpressionSet(assayData=as.matrix(dDat[,oInd]))
  pData(oDat) <- oPheno
 
  myDes1c <- model.matrix(~plate + TotalReads + age + case*stimulation, eDat)
  myDesyc <- model.matrix(~plate + TotalReads + case*stimulation, yDat)
  myDesoc <- model.matrix(~plate + TotalReads + case*stimulation, oDat)
 
  v <- voom(exprs(eDat), design=myDes1c, plot=FALSE, lib.size=libNormc)
  vy <- voom(yDat, design=myDesyc, plot=FALSE, lib.size=ylibNorm)
  vo <- voom(oDat, design=myDesoc, plot=FALSE, lib.size=olibNorm)
 
  coefsc <- colnames(myDes1c)[9:12]
  labsc <- c("DMSO vs csp", "DMSO vs hbs", "DMSO vs csp:disease", "DMSO vs hbs:disease")
  agey <- c("both", "young", "old")
  
  if(FALSE) {
    
      ## age == both model
      ranCorc <- duplicateCorrelation(v, design=myDes1c,
                                     block=dPheno$SID)$consensus.correlation
      fit1c <- lmFit(v, myDes1c, block=dPheno$SID, correlation=ranCorc)
      save(fit1c, ranCorc, file=paste0(ROOT, "data/lmFitDataWithCoVarM3caseNoAMA.Rda"))
     
     ## run young age model
     ranCorYoungc <- duplicateCorrelation(vy, design=myDesyc,
                                     block=yPheno$SID)$consensus.correlation
     fitYoungc <- lmFit(vy, myDesyc, block=yPheno$SID, correlation=ranCorYoungc)
     save(fitYoungc, ranCorYoungc, file=paste0(ROOT, "data/lmFitDataWithCoVarYoungcaseNoAMA.Rda"))
    
     ## run old age model
     ranCorOldc <- duplicateCorrelation(vo, design=myDesoc,
                                     block=oPheno$SID)$consensus.correlation
     fitOldc <- lmFit(vo, myDesoc, block=oPheno$SID, correlation=ranCorOldc)
     save(fitOldc, ranCorOldc, file=paste0(ROOT, "data/lmFitDataWithCoVarOldcaseNoAMA.Rda"))
     
  } else {
     load(file=paste0(ROOT, "data/lmFitDataWithCoVarM3caseNoAMA.Rda"))    ## load fit1
     load(file=paste0(ROOT, "data/lmFitDataWithCoVarYoungcaseNoAMA.Rda")) ## load fitYoung
     load(file=paste0(ROOT, "data/lmFitDataWithCoVarOldcaseNoAMA.Rda"))   ## load fitOld
  } ## end if FALSE
  
   ## assemble various fits
   allFitc <- list(fit1c, fitYoungc, fitOldc)
  
  ## extract data.tables and graphs from all fits
  allDTc <- allDEGc <- allGraphsc <- selectDEG <- list()
  ageInd <- 0
  for(fit in allFitc) {
      
      ## set age for this fit
      ageInd <- ageInd + 1
      ageLab <- agey[ageInd]
      
      ## shrinkage
      fit2 <- eBayes(fit, trend=FALSE)
    
      ## cycle through coefficients
      colInd <- c(1,4,5)
      DEG <- NULL
      for(i in 1:length(coefsc)) {
           
           c <- coefsc[i]
           labb <- paste(labsc[i], ageLab, sep=": ")
    
           ## generate datatables
           tmp <- topTable(fit2, number=Inf, coef=c, sort="P")
           ##write.csv(tmp, file=paste0(ROOT, "data/", labs[i], "Result.csv"), row.names=TRUE)
           indy <- tmp$P.Value <= pCut
           tmpy <-  signif(tmp[indy, colInd], 3)
           colnames(tmpy) <- c("log FC",  "P.value", "FDR")
           allDTc[[labb]] <- tmpy
           DEG <- c(DEG, sum(tmpy$FDR <= FDRCut))
   
           ## generate graphs
          graphSet <- list()
          ress1 <- tmp
          ress1$FDR <- ifelse(ress1$adj.P.Val > FDRCut, "NonSig", "Sig")   ## label significant genes
          
          graphSet[[1]] <- ggplot(ress1, aes(x=P.Value)) + geom_histogram(binwidth=.05) +
            xlab("p-value") + ggtitle(label=labb) + theme_bw() + theme(aspect.ratio=1,
                                                                       axis.text.x=element_text(size=10),
                                                plot.title = element_text(size = 15)) + 
            scale_x_continuous(breaks=c(0, 0.5, 1))
          
          graphSet[[2]] <- ggplot(ress1, aes(x=logFC, y=-log10(P.Value), color=FDR)) +
            geom_point(shape=1) + ggtitle("logFC vs -log10 p-value") + theme_bw() + 
            theme(axis.title.y=element_text(size=10), axis.text.x=element_text(size=6), 
                  plot.title = element_text(size = 13)) + 
            geom_vline(xintercept=c(-1,1), linetype="dashed", color="blue") + 
          geom_hline(yintercept=-log10(c(0.01)), linetype="dashed", color="blue") +
            theme(legend.position="none") +  scale_colour_manual(values=c("black", "red")) +
            theme(aspect.ratio=1)
          
          graphSet[[3]] <- ggplot(ress1, aes(x=adj.P.Val)) + geom_histogram(binwidth=.05) +
            xlab("p-value") + ggtitle("FDR") + theme_bw() + 
            theme(aspect.ratio=1, plot.title = element_text(size = 15)) + 
            scale_x_continuous(breaks=c(0, 0.5, 1))
          
          allGraphsc[[labb]] <- graphSet
          
          ## get index label for vaccine model
          selectDEGC <- coefs[i]
          selectDEGLab <-  paste(labs[i], ageLab, sep=": ")
          selectDEGDT <- allCompDT[[selectDEGLab]]
          
          ## get disease results for significant gene sets with vaccine model
          if(is.null(dim(selectDEGDT))) {
             selectDEG[[labb]] <- -1            
          } else {
              tmp2 <-  signif(tmp[rownames(selectDEGDT),colInd], 3)
              
              ## recalculate FDRS for just the previously selected genes
              fdrs <- p.adjust(tmp2$P.Value, method="BH")
              tmp2$adj.P.Val <- signif(fdrs, 3)
              tmp2 <- tmp2[order(tmp2$P.Value),]
              selectDEG[[labb]] <- tmp2
          }
    
      }
    
      allDEGc <- rbind(allDEGc, DEG)
   
  } ## end for fit
  
  ## reorder the lists for better presentation
  seqq <- c(1,5,9)
  newOrder <- names(allGraphsc)[c(seqq, seqq+1, seqq+2, seqq+3)]
  allGraphsc <- allGraphsc[newOrder]
  allDTc <- allDTc[newOrder]
  selectDEG <- selectDEG[newOrder]
  
  ## show only results for interactions
  printGraphsc <- allGraphsc[7:12]
  printDTc <- allDTc[7:12]
  printSelectDEG <- selectDEG[7:12]
  printDEGTabc <- allDEGc[,3:4]
  dimnames(printDEGTabc) <- list(agey, c("csp", "hbs"))
    
  ## assemble output in string format for knit_child  
  outCaseDEGc = NULL
  for (i in 1:length(printGraphsc)) {
    
      ## vaccine model
      outCaseDEGc <- c(outCaseDEGc,  paste0("##`r names(printDT)[",i, "]` \n"))
      outCaseDEGc <- c(outCaseDEGc,  paste0("```{r results='asis', echo=FALSE}\ngrid.arrange(printGraphs[[", i, "]][[1]], printGraphs[[", i, "]][[2]], printGraphs[[", i, "]][[3]], nrow=1, ncol=3)\n```\n"))
    
      outCaseDEGc <- c(outCaseDEGc,  paste0("##`r names(printDT)[",i, "]` {.smaller}\n"))
     if(is.null(dim(printDT[[i]]))) {
           knit_expanded <- paste0("No significant genes found\n\n <br>") 
      }else {
        knit_expanded <- paste0("```{r results='asis', echo=FALSE}\n\ndatatable(printDT[[", i, "]])\n\n```\n")
      }
      outCaseDEGc <-  c(outCaseDEGc, knit_expanded)
    
      ## disease model for significant genes from vaccine model
       outCaseDEGc <- c(outCaseDEGc,  paste0("##`r names(printDTc)[",i, "]`| disease model for significant vaccine effects {.smaller}\n"))
       if(is.null(dim(printSelectDEG[[i]]))) {
           knit_expanded <- paste0("No significant genes found\n\n <br>") 
      }else {
        knit_expanded <- paste0("```{r results='asis', echo=FALSE}\n\ndatatable(printSelectDEG[[", i, "]])\n\n```\n")
      }
     outCaseDEGc = c(outCaseDEGc, knit_expanded)
      
      ## disease model
     outCaseDEGc <- c(outCaseDEGc,  paste0("##`r names(printDTc)[",i, "]` \n"))
     outCaseDEGc <- c(outCaseDEGc,  paste0("```{r results='asis', echo=FALSE}\ngrid.arrange(printGraphsc[[", i, "]][[1]], printGraphsc[[", i, "]][[2]], printGraphsc[[", i, "]][[3]], nrow=1, ncol=3)\n```\n"))
    
     outCaseDEGc <- c(outCaseDEGc,  paste0("##`r names(printDTc)[",i, "]` {.smaller}\n"))
     if(is.null(dim(printDTc[[i]]))) {
           knit_expanded <- paste0("No significant genes found\n\n <br>") 
      }else {
        knit_expanded <- paste0("```{r results='asis', echo=FALSE}\n\ndatatable(printDTc[[", i, "]])\n\n```\n")
      }
     outCaseDEGc = c(outCaseDEGc, knit_expanded)
  } ## end for i
  
  ##<!--Output DEG results results-->
  ##<!--`r paste(knit_child(text = outCaseDEGc), collapse = '\n')` -->


  
```

<!--Output DEG results results-->
`r paste(knit_child(text = outCaseDEGc), collapse = '\n')`

##DEG summary stimulation vaccine interaction | FDR cutoff: `r FDRCut`

Number of significant genes for dmso versus stimulants and vaccine interaction.

```{r results='asis', echo=FALSE}
   print(xtable(printDEGTab, align=rep("c", ncol(printDEGTab)+1), auto=TRUE), type = "html")
```

##DEG summary stimulation disease interaction | FDR cutoff: `r FDRCut`

Number of significant genes for dmso versus stimulants and vaccine interaction.

```{r results='asis', echo=FALSE}
   print(xtable(printDEGTabc, align=rep("c", ncol(printDEGTabc)+1), auto=TRUE), type = "html")
```

##SessionInfo{.samller}
```{r sessionInfo}

  ## generate version information
   sessionInfo()
```
  